Investigating the influence of genre preferences on movie/series rating patterns

### 2024-03-25 gen/paper/paper.Rmd
paper knits

---
title: Investigating the influence of genre preferences on movie/series rating patterns
  over time
author: "Team 1 - Timo Philipse, Bram Teunissen, Rodrigo Pačeko Rudzājs, Dian de Ridder & Linda van den Boogaart"
output: pdf_document
editor_options:
  markdown:
    wrap: 72
---

@@

The research results shed light on how consumers use dynamic genres to
shape their satisfaction with a movie or TV series. The findings suggest
that some genres are preferred in certain years over others. For
example, war movies tend to receive higher ratings during world wars
compared to times of peace. This implies that consumers use these
productions as references to rate them in context of the sign of the
times. Managers may want to consider adjusting their productions to
better suit the elements in those preferred years.

The study also found that recent fantasy movies tend to receive lower
ratings from consumers, while those produced in the 70s are rated more
favorably. Overall, the results suggest that managers may face
challenges in meeting previously set expectations and that these
findings provide a reference frame for when those expectations are met.

Additionally, the results also provide insights into how genres are
related to average ratings across another. Currently, every production
has a similar pattern of ingredients. Most movies contain hints of drama
and romance. Whether it is in a distropian sci-fi survival movie or an
horror action series. However, the results indicated that these common
genres are not significantly related to the average ratings. In contrast
some movies and series do contain unique elements do results in
significant relationships. For example, biographies and animations. This
may imply that consumers may not rely so heavily on past genres that are
common because they gradually change with times without consumers
noticing. (Think of it like a birthday party with your family when you
are 12. Your aunt keeps telling you "How much you have grown!", but you
don't notice any difference) In contrast, the uniqueness of biographies
does sharply stand out. Therefore, managers may want to focus more
strongly on the less common genre features compared to more common ones.

@@

In this paper we investigate consumers’ movie and series preferences
using four of IMDb's datasets containing movie data. First, these

datasets are loaded, cleaned, inspected, merged and cleaned. Next, the
data is analyzed using a regression analysis. This regression analysis
regresses all movie and series genre's against the average rating's over
a sliding period of 5 years. Additionally, the variables title,
movieType and time are used as control variables to improve the
interpret ability of the estimates.

@@

### Descriptives
The data set contains of 70087 films, of which 60246 are movies and 9841 are series.  These films have on average a rating of 6.3, with 10 being the highest and 1 being the lowest. After removing the votes to all being above 100, the average number of votes on a film became 7146. When looking at the years, the most amount of films were released on IMDb in 2019, with 2586 films. Also when comparing the 5 year range of films released,2015 to 2019 had the most films released, with 11353 films. Looking at the runtime of the films, on average a film is 101 minutes long.

@@

@@

A point of concern, however, is that these average ratings are not
measured over time but are a fixed sum up to and including the data of
analysis. This means that if series fall out of favor and their rating
drops this remains unobservant. Given that the continuation of a series
is determined by its success (a second season is not produced if the
first one was a flop) the following assumption can be made: if a genre
is not preferred throughout the continuance of a series, the rating is
low in the starting year and thus unlikely to be significant.
Additionally, since IMDb was not around in 1925, average ratings for
movies and series originating from that time may be strongly biased
because a niche selection of the population may be watching these movies
in the first place.

@@

Last, to see how preferences change over time we build on the assumption
that if a movie or series is a success they gain many positive ratings
in the beginning and less so over time. As a result, there is a peak
during (the first season) release allowing us to only use the startYear
as a means to calculate behavior over time. Yet, the individual years
are unlikely to have a remarkable effect on the genre preference trends.
This is because the effects are so nuanced that they are unlikely to be
detected. Therefore, a five-year window was created that captures all 
genre preferences for that period of time.

@@

@@

knitr::include_graphics("../../gen/data_preparation/qq_normality.png")

```

@@

```{r echo=FALSE}
knitr::include_graphics("../../gen/data_preparation/homoskedacity.png")
```
```{r echo=FALSE}
knitr::include_graphics("../../gen/data_preparation/independence.png")
```

@@

This research illustrates that genre preferences may be related to the
average IMDb ratings and vary over time. While further research is
necessary to fine-tune these claims the current research does provide
insight into how raters' watch patterns appear to alter. For example, it
shows that the common ingredients like romance and action may not stand
out as much as biography and history might. 
The results of this analysis can be, at best, interpreted as correlational 
given the vast amount of assumptions and potential hidden dangers of 
sampling biases (e.g., film-noir and war genres). While this has been 
partially mitigated by loosening some of the underlying assumptions and 
adding fixed effects, the fact remains that these appear as trend 
observations. Therefore, the results do not necessarily conclude that some 
genres lead to higher average ratings. However, the results do indicate that 
there may be preferences over time that may lead to higher average ratings.

@@

```{r saving}
library(rmarkdown)

render(input = "../../gen/paper/paper.Rmd",
       output_format = "pdf_document",
       output_file = "../../gen/paper/paper.pdf")
```

>>>>

---
title: Investigating the influence of genre preferences on movie/series rating patterns
  over time
author: "Team 1 - Timo Philipse, Bram Teunissen, Rodrigo Pačeko Rudzājs, Dian de Ridder & Linda van den Boogaart"
output: 
  pdf_document: {}
editor_options:
  markdown:
    wrap: 72
---

@@

The research results shed light on how consumers use dynamic genres to
shape their satisfaction with a movie or TV series. The findings suggest
that some genres are preferred in certain years over others. For
example, war movies tend to receive higher ratings during world wars
compared to times of peace. This may imply that consumers use these
productions as references to rate them in context of the sign of the
times. Managers may want to consider adjusting their productions to
better suit the elements in those preferred years.

The study also found that recent fantasy movies tend to receive lower
ratings from consumers, while those produced in the 70s are rated more
favorably. Overall, the results may suggest that managers may face
challenges in meeting previously set expectations and that these
findings provide a reference frame for when those expectations are met.

Additionally, the results also provide insights into how genres are
related to average ratings across another. Currently, every production
has a similar pattern of ingredients. Most movies contain hints of drama
and romance, regardless of whether it is in a distropian sci-fi survival movie or an horror action series. However, the results indicated that these common genres are not significantly related to the average ratings. In contrast some movies and series do contain unique elements do results in
significant relationships. For example, biographies and animations. This
may imply that consumers may not rely so heavily on past genres that are
common because they gradually change with times without consumers
noticing. (Think of it like a birthday party with your family when you
are 12. Your aunt keeps telling you "How much you have grown!", but you
don't notice any difference) In contrast, the uniqueness of biographies
does sharply stand out. Therefore, managers may want to focus more
strongly on the less common genre features compared to more common ones.

@@

In this paper we investigate consumers’ movie and series preferences
using four of IMDb's data sets containing movie data. First, these

Data sets are loaded, cleaned, inspected, merged and cleaned. Next, the
data is analyzed using a regression analysis. This method was chosen due to the continuous nature of the dependent variable and the categorical nature of the independent variables. The reason why a regression was chosen over the alternative of an ANOVA was due to the size and complexity of the analysis. There are 27 genres included as a dependent variable, and are measure over time. To include these features into an ANOVA is far more complex compared to a regression analysis.

This regression analysis regresses all movie and series genre's against the average rating's over a sliding period of 5 years. Additionally, the variables title, movieType and time are used as control variables to improve the interpret ability of the estimates. Additionally, because some of the assumptions were only robustly satisfied, the standard errors were clustered. The analysis resulted in a 20 window output, one for each year window.

A note of attention should be, that despite the richness of the data, the interpretation of the results remain descriptive at bests. This is because the average movie ratings are not reported over time but are merely a static measure of today's average. This implies that a single coefficient is read as "The genre X produced in the year Y is positively/negatively associated with the average movie rating of today."

@@

### Descriptives

The data set contains of 70087 films, of which 60246 are movies and 9841 are series.  These films have on average a rating of 6.3, with 10 being the highest and 1 being the lowest. After removing the votes to all being above 100, the average number of votes on a film became 7146. When looking at the years, the most amount of films were released on IMDb in 2019, with 2586 films. Also when comparing the 5 year range of films released,2015 to 2019 had the most films released, with 11353 films. LAst, when looking at the run time of the films, on average a film is 101 minutes long.

@@

Please note the average movie rating is a static measure that does not change the dynamics of the rating. For example, we can only see today's average rating but not that of 2 hours, 2 days, or even 2 years ago.

@@

A point of concern, however, is that these average ratings are not
measured over time but are a fixed sum up to and including the data of
analysis. This means that if series fall out of favor and their rating
drops this remains unobservant. Given that the continuation of a series
is determined by its success (a second season is not produced if the
first one was a flop) the following assumption can be made: if a genre
is not preferred throughout the continuance of a series, the rating is
low in the starting year and thus unlikely to be significant.
Additionally, since IMDb was not around in 1925, average ratings for
movies and series originating from that time may be strongly biased
because a niche selection of the population may be watching these movies
in the first place. Therefore, regardless of the additions of these fixed effects the results of the analysis are correlational at best.

@@

Last, to see how preferences change over time we build on the assumption
that if a movie or series is a success they gain many positive ratings
in the beginning and less so over time. As a result, there is a peak
during (the first season) release allowing us to only use the startYear
as a means to calculate behavior over time. Yet, the individual years
are unlikely to have a remarkable effect on the genre preference trends.
This is because the effects are so nuanced that they are unlikely to be
detected. Therefore, a five-year window was created that captures all 
genre preferences for that period of time. This window was constructed by grouping all productions in an interval of five years.

@@

#### Normality

@@

knitr::include_graphics("../../gen/data_preparation/qq_normality.png")
```

@@

#### Homoskedacity

```{r echo=FALSE}
knitr::include_graphics("../../gen/data_preparation/homoskedacity.png")
```

#### Independence

```{r echo=FALSE}
knitr::include_graphics("../../gen/data_preparation/independence.png")
```

#### Linearity

@@

This research illustrates that genre preferences may be related to the
average IMDb ratings and vary over time. While further research is
necessary to fine-tune these claims the current research does provide
insight into how raters' watch patterns appear to alter. For example, it
shows that the common ingredients like romance and action may not stand
out as much as biography and history might.

The results of this analysis can be, at best, interpreted as descriptive 
given the vast amount of assumptions and potential hidden dangers of 
sampling biases (e.g., film-noir and war genres). While this has been 
partially mitigated by loosening some of the underlying assumptions (clustering standard errors and selecting a better heterogeneity method) and adding fixed effects, the fact remains that these appear as trend 
observations. Therefore, the results do not necessarily conclude that some 
genres lead to higher average ratings. However, the results do indicate that there may be preferences over time that may lead to higher average ratings.

@@
